---
title: "Week8_Assignment- Computational Techniques in Data Science"
author: "Jackline Mboya - 193670"
date: "2025-05-12"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r setup-mirror, include=FALSE}
options(repos = c(CRAN = "https://cloud.r-project.org"))

```



```{r}
install.packages("Benchmarking")
```

# Question 1: Hospital Department Efficiency
Question:

You are analyzing the performance of 4 hospital departments. Each department has a different combination of medical staff and beds. You want to know:

Which department is most efficient in treating patients with limited resources?

### Inputs:
Doctors: [40, 35, 45, 38]

Nurses: [80, 70, 85, 75]

Beds: [150, 140, 160, 145]

### Outputs:
Patients treated:[2000, 1900, 2100, 1950]

Survival rate: [0.97, 0.96, 0.98, 0.95]

```{r}
library(Benchmarking)
```


```{r}
# Inputs
X_hospital <- matrix(c(
  40, 80, 150,
  35, 70, 140,
  45, 85, 160,
  38, 75, 145
), ncol = 3, byrow = TRUE)

# Outputs
Y_hospital <- matrix(c(
  2000, 0.97,
  1900, 0.96,
  2100, 0.98,
  1950, 0.95
), ncol = 2, byrow = TRUE)

# Run CCR Input-Oriented DEA
eff_hospital <- dea(X_hospital, Y_hospital, RTS = "crs", ORIENTATION = "in")

# Print results
cat("Hospital Department Efficiencies:\n")
print(round(eff_hospital$eff, 4))

# Identify most efficient hospital
most_efficient_hospital <- which.max(eff_hospital$eff)
cat("\nThe most efficient hospital department is: Hospital", LETTERS[most_efficient_hospital], "\n")
```
# Question 2: Retail Store Efficiency
Question:

A retail company wants to assess whether its 4 branches are efficiently turning staff and marketing investment into profit.

Which store branch is the best performer given similar resource levels?

### Inputs:
Marketing Spend (USD): [$12,000, $10,000, $11,500, $9,800]

Employees: [20, 18, 22, 19]

### Outputs:
Revenue (USD): [$110,000, $100,000, $120,000, $95,000]

Customers: [2,300, 2,000, 2,600, 1,900]

```{r}
# Inputs
X_retail <- matrix(c(
  12000, 20,
  10000, 18,
  11500, 22,
  9800, 19
), ncol = 2, byrow = TRUE)

# Outputs
Y_retail <- matrix(c(
  110000, 2300,
  100000, 2000,
  120000, 2600,
  95000, 1900
), ncol = 2, byrow = TRUE)

# Run CCR Input-Oriented DEA
eff_retail <- dea(X_retail, Y_retail, RTS = "crs", ORIENTATION = "in")

# Print results
cat("Retail Store Efficiencies:\n")
print(round(eff_retail$eff, 4))

# Identify most efficient store
most_efficient_store <- which.max(eff_retail$eff)
cat("\nThe most efficient retail store is: Store", LETTERS[most_efficient_store], "\n")
```
# Question 3: School Performance Evaluation
Question:

A school district monitors 4 schools to find out which one is using its resources most efficiently to produce better academic results.

Which school is most efficient in converting resources into academic outcomes?

### Inputs:
Teachers: 30, 25, 20, 28

Budget (USD): [$800,000, $750,000, $700,000, $770,000]

### Outputs:
Graduation Rate (%): [85, 90, 82, 88]

Average Test Score (out of 100): [78, 80, 76, 79]

```{r}
# Inputs
X_school <- matrix(c(
  30, 800000,
  25, 750000,
  20, 700000,
  28, 770000
), ncol = 2, byrow = TRUE)

# Outputs
Y_school <- matrix(c(
  85, 78,
  90, 80,
  82, 76,
  88, 79
), ncol = 2, byrow = TRUE)

# Run CCR Input-Oriented DEA
eff_school <- dea(X_school, Y_school, RTS = "crs", ORIENTATION = "in")

# Print results
cat("School Performance Efficiencies:\n")
print(round(eff_school$eff, 4))

# Identify most efficient school
most_efficient_school <- which.max(eff_school$eff)
cat("\nThe most efficient school is: School", LETTERS[most_efficient_school], "\n")
```

```{r}

```

